The Unicode Cookbook \linebreak for Linguists

### 2015-11-13 book/chapters/ipa_meets_unicode.tex
adding note on tone and stress markers

Although linguists might expect these characters to belong together with the
character in front of them, at least for <ʰ> \textsc{modifier letter small h} at
\uni{02B0} the Unicode Consortium's decision to treat it as a separate character
is also linguistically correct, because according to the IPA it can be used both
for post-aspiration (following the base character) and pre-aspiration (preceding
the base character). The default combination of Spacing Modifiers with the
preceding character can be specified in orthography profiles (see
Chapter~\ref{orthography-profiles}). Note that there is an alternative 
homoglyph, namely the \textsc{combining latin small letter h} at \uni{036A}. We 
propose not to use this combining diacritical mark, but to standardize on 
Unicode modifier letters for all superscript letters in IPA.

>>>>

Although linguists might expect these characters to belong together with the
character in front of them, at least for tone letters, stress symbols and <ʰ>
\textsc{modifier letter small h} at \uni{02B0} the Unicode Consortium's decision
to treat it as a separate character is also linguistically correct.

   \item According to the IPA, <ʰ> can be used both as <\dia{02B0}> for
         post-aspiration (following the base character) and as
         <\diareverse{02B0}> for pre-aspiration (preceding the base character).
         Note that there is yet another homoglyph, namely the
         \textsc{combining latin small letter h} at \uni{036A}. We propose not
         to use this combining diacritical mark, but to standardize on Unicode
         modifier letters for all superscript letters in IPA.

   \item Tone letters are normally written at the end of the syllable,
         possibly occurring immediately adjacent to a consonant in the coda of
         the syllable. Such tone markers should of course not be treated as
         belonging to this consonant, so we propose to treat tone letters as 
         separate segments.

   \item Stress markers <\diareverse{02C8}> at \uni{02C8} and
         <\diareverse{02CC}> at \uni{02CC} have a very similar
         distribution in that they normally are written at the start of
         the stressed syllable. In a sense, they thus belong to the
         characters following the stress marker, but it would be wrong to
         cluster them together with whatever segment is at the start of
         the syllable. So, like tone letters, we propose to treat stress
         markers as separate segments.

If intended, then any default combination of Spacing Modifiers
with the preceding character can be specified in orthography profiles (see
Chapter~\ref{orthography-profiles}).
